Poster Data Analysis

### 2018-09-12 notebooks/poster_analysis.Rmd
added dendrogram for heatmap data

Please press the "code" button for each chunk to view code.
```{r message=FALSE, warning=FALSE}
# Libraries
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr, ggplot2, tidyr, cowplot, viridis, forcats, radiant.data, purrr)
```

@@

## Heatmap plot
This plot represents the individual resistance mechanisms (x-axis) present in each isolate (y-axis), for each animal species (colors). Grey squares represent genes/mutations that ARIBA flagged as bad quality in their report. Dark color indicates presence, while the light color indicates absence.
```{r fig.height=30, fig.width=10, message=FALSE, warning=FALSE}
megares_long <- megares_results %>%
  gather(gene, value, -c(id, type))

@@

megares_long %>%
  rbind(resfinder_long) %>%
  mutate(type2 = ifelse(type == "mut", 1, 2),
         value = factor(ifelse(is.na(value) == TRUE, 2, value))) %>%
  spread(id, value, fill = 0) %>%
  gather(key = "id", value = "result", -c(gene, type, type2)) %>%
  left_join(species_data, by = "id") %>%
  mutate(result = ifelse(result == 2, NA, result),
         fillcol = paste(species, result, sep = "_"),
         fillcol = ifelse(is.na(result) == TRUE, NA, fillcol)) %>%
  ggplot(aes(fct_reorder(gene, type2),
             factor(id,
                    levels = unique(.[order(.$species,.$id),"id",drop = TRUE])),
             fill = fillcol,
             alpha = result)) +
  geom_tile(color = "white")+
  geom_vline(xintercept = 14.5)+
  scale_fill_brewer(palette = "Paired",
                    na.value = "#6b6767")+
  theme_minimal()+
  labs(fill = NULL)+
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1,
                                   vjust = 0.3,
                                   size = 7),
        axis.title = element_blank(),
        axis.text.y = element_blank(),
        panel.grid = element_blank(),
        legend.position = "right")+
  guides(alpha = FALSE)+
  coord_fixed()

>>>>

Please press the "code" button for each chunk to view code.
```{r message=FALSE, warning=FALSE}
# Libraries
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr, ggplot2, tidyr, cowplot, viridis, forcats, radiant.data, purrr, phangorn, gridExtra, ggtree)
```

@@

## Heatmap plot
This plot represents the individual resistance mechanisms (x-axis) present in each isolate (y-axis), for each animal species (colors). Grey squares represent genes/mutations that ARIBA flagged as bad quality in their report. Dark color indicates presence, while the light color indicates absence.
```{r fig.height=30, fig.width=10, message=FALSE, warning=FALSE}
# Adjust MLST data frame
sequence_types_adjusted <- mlst_results %>%
  mutate_all(funs(gsub("\\*", "", .))) %>%
  select(-ST)

row.names(sequence_types_adjusted) <- sequence_types_adjusted$id

mlst <- sequence_types_adjusted[,-1]

# Calculate distance matrix from sequence typing alleles
d.mlst.distances <- matrix(0, ncol=nrow(mlst), nrow=nrow(mlst))
for (i in 1:(nrow(mlst)-1)) {
  for (j in (i+1):nrow(mlst)){
    d.mlst.distances[i,j] <- sum(mlst[i, ] != mlst[j, ])
    d.mlst.distances[j,i] <- sum(mlst[i, ] != mlst[j, ])
  }
}

# Create tree from distance matrix
tree <- upgma(d.mlst.distances)
id_label_tree <- tree 
id_label_tree$tip.label <- sequence_types_adjusted$id
tree$tip.label <- mlst_results$ST

# Create dummy plot of id_label_tree to extract correct order of ID's
p0 <- ggtree(id_label_tree)+
  geom_treescale()+
  geom_tiplab(align = TRUE, linesize = 0, hjust = -0.2, size = 1.5)+
  xlim(0,3.6)

# Get plot build data
p1b <- ggplot_build(p0)

# Save plot of tree
p1 <- ggtree(tree)+
  geom_treescale()+
  geom_tiplab(align = TRUE, linesize = 0, hjust = -0.2, size = 1.5)+
  xlim(0,3.6)

# Create heatmap data frame from gene data
megares_long <- megares_results %>%
  gather(gene, value, -c(id, type))

@@

heatmap_data <- megares_long %>%
  rbind(resfinder_long) %>%
  mutate(type2 = ifelse(type == "mut", 1, 2),
         value = factor(ifelse(is.na(value) == TRUE, 2, value))) %>%
  spread(id, value, fill = 0) %>%
  gather(key = "id", value = "result", -c(gene, type, type2)) %>%
  left_join(species_data, by = "id") %>%
  mutate(result = ifelse(result == 2, NA, result),
         fillcol = paste(species, result, sep = "_"),
         fillcol = ifelse(is.na(result) == TRUE, NA, fillcol))

# Arrange id-column after tip-labels in tree
heatmap_data <- heatmap_data %>%
  mutate(id = factor(id, levels = p1b$data[[5]] %>% arrange(y) %>% pull(label)))

# Plot the heatmap
p2 <- ggplot(heatmap_data, aes(fct_reorder(gene, type2),
             factor(id),
             fill = fillcol,
             alpha = result)) +
  geom_tile(color = "white")+
  geom_vline(xintercept = 13.5)+
  scale_fill_brewer(palette = "Paired",
                    na.value = "#6b6767")+
  theme_minimal()+
  labs(fill = NULL)+
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1,
                                   vjust = 0.3,
                                   size = 7),
        axis.title = element_blank(),
        axis.text.y = element_blank(),
        panel.grid = element_blank(),
        legend.position = "right")+
  guides(alpha = FALSE)+
  coord_fixed()

# Arrange both plots into one figure
grid.arrange(arrangeGrob(p1 + theme(plot.margin=margin(68,-10,149,0)),
                         heights = c(1.9,-0.097)),
             arrangeGrob(p2 + theme(axis.text.y=element_blank(),
                                    plot.margin = margin(0,0,0,-8)),
                         heights=c(0.98,-0.03)), 
             ncol=2)
